### 2009-06-04 response-to-reviews2.txt

We hope & anticipate that the sophistication of the constraint schemes
used with these sorts of model will steadily increase over
time. Eventually we hope to develop alignment constraints,
simultaneous alignment-and-fold constraints, and other constraints
(and/or sampling schemes) to enable efficient analysis of multiple
phylogenetically-related sequences. The present work represents a
first attempt at using these sorts of models for more than two
sequences, but we hope it will not be the last attempt, and we
acknowledge room for improvement.

> It is still not clear eenough what the authors' model is, though it is 
> supposed to be a main result. Page 7 states "a prescriptive formalism 
> for transducers is given in Text S1". Relegating the deliverables of a 
> manuscript to supplementary material is not an appropriate use of 
> supplementary material; a reader should not have to refer to a second 
> manuscript to understand the main points. The authors need to find a 
> way to explain their main results in the actual manuscript. There is 
> still much vagueness and redundancy in this draft, which could be 
> removed and replaced with more relevant explanation. For example, in 
> this same section where results are relegated to supplementary 
> material, there's a whole redundant paragraph (p.8) on how conditional 
> models represent directed edges; but this concept has already been 
> explained sufficiently. This is followed (bottom p. 8, "Consider a 
> phylogenetic tree...") by yet another statement of the same concept of 
> using conditional models, still in
> terms too vague and conceptual to enable reproduction of the work.

Thank you for pointing this out.  We have attempted to make the main
text a self-contained description of the composition algorithm,
including its key corollaries (elimination of null states and DP
inference).  The presentation of the composition algorithm in the
supplement (text S1) now merely reprises the presentation in the main
text using more formal notation. Likewise, the null cycle elimination
algorithm is discussed verbally in the main paper, and described again
using formal notation in text S2. Texts S3 and S4 describe aspects of
the software that are also probably too detailed for the main paper.

We hope this is an appropriate use of the supplementary information,
but we welcome additional suggestions to improve the flow of the
manuscript for the reader.

@@

> p.13 The DP inference algorithm section says "the standard CYK ... 
> algorithms fail". But then it says "Maximum likelihood inference is 
> performed with a CYK algorithm", and even "maximum likelihood 
> inference can therefore be performed with a standard three sequence 
> CYK algorithm". Which is it?

Thank you for pointing out this inconsistency, which we have clarified
in the revised text.  Using the standard CYK algorithm on the composed
model fails; however, the model can be simply modified as described in
the revised text to permit standard CYK to work properly.

> p.14 It seems possible that the problem with "null cycles" (residues 
> in ancestors that leave no homologous residues in observed sequences) 
> may simply go away if you view the problem as a missing data problem, 
> and that the likelihood being inferred is the likelihood of alignment 
> columns needed to account for the observed data (marginalized over 
> missing columns needed to account for null cycles) rather than the 
> joint likelihood of the "observed" columns and the "null cycle" 
> columns. This is what dropped out of the equations in the phylogenetic 
> inference described by Rivas and Eddy (2008) PLoS CB 4:e1000172, for 
> example.

We completely agree. Such an algorithm was initially a goal of this
work, but was dropped. We thank the reviewer for challenging us to
try again; a verbal presentation of the algorithm we came up with is
given in the revised text, with a full formal description (including
notation) in supplement S2. (Since this is a standalone result that
can be applied to any SCFG, we think it fits better as a technical
appendix in the supplement, with a verbal description in the main
paper; the obvious cost is omitting some detail from the main paper.
We welcome advice on how to improve the flow generally.)

As a technical aside, the primary difficulty with SCFG null cycle
elimination is that while this marginalization is reasonably
straightforward for independent-sites models and HMMs, essentially
involving summing a geometric series to obtain a correction factor,
the possible presence of null bifurcations complicates matters for
SCFGs.  Indeed, no *closed form* elimination is possible (because of
bifurcations: such an elimination would involve solving simultaneous
nonlinear equations, as shown in supplement S2) but an iterative
approximation (e.g. Newton-Raphson) appears to be adequate.

@@

> p.21 in general the whole "reconstructing RNAs with a simple model" 
> section is too chatty and imprecise. "Both predicted ancestral 
> structures closely match the RFAM-annotated consensus structures", for 
> example, is not a quantitative result, and was not set up to address 
> any particular focused question (like, for instance, does our method 
> do better than some simpler approach of addressing the same problem?) 
> Again too much space is devoted to vague forward-looking conceptual 
> claims ("Though it's beyond the scope of the paper, a revised pairwise 
> model could readily be designed...") and too little attention is given 
> to the actual model and problems at hand.

We have reworded this section per our new focus on the simulation results.

> Table 2: it's unclear why we need separate states L and I_L, or S and 
> I_S; these nonterminals have identical production rules.

This is a case of shoe-horning the grammar into the input format
accepted by indiegram (and by our composition algorithm).

### 2009-06-04 paper.tex
renamed software appendix (Text S3) to Text S4; added new Text S3 for TKFST grammars; added a bunch of other stuff

\supptext{2} includes a complete formal algorithm for exact null-cycle elimination in SCFGs,
along with procedures for probabilistically restoring null cycles
to sampled parse trees and Inside-Outside expectation counts.
Informally, the essence of the algorithm is contained within the following two steps:
\begin{description}
\item[(i)] separating bifurcations
into those which have one or more empty children (and can therefore be
represented using transition or termination rules) and those that have
two nonempty children;
\item[(ii)] replacing all empty paths through null states with effective direct transitions between non-null states,
obtaining sum-over-paths probabilities by inverting the grammar's transition matrix.
\end{description}

@@

The TKF Structure Tree (TKFST) model is based on the
Thorne-Kishino-Felsenstein (TKF) model of the stochastic evolution of
primary sequences via indel events \cite{ThorneEtal91}.  In the
original TKF model, sequence evolves under a time-homogeneous linear
birth-death-immigration process \cite{Kendall1948}.  Single characters
are inserted with rate $\lambda$ and deleted with rate $\mu$.  At
equilibrium, sequences obey a geometric length distribution with
parameter $\kappa$.  Although this model has flaws (e.g., it lacks
affine gap penalties, rate heterogeneity and context-dependent
mutation rates), it illustrates many of the key ideas used by more
sophisticated indel models, notably the possibility for systematic
derivation of pairwise alignment automata from first principles via
analysis of birth-death processes \cite{Feller71,ThorneEtal91}.

The TKF Structure Tree model is an extension of the TKF model to RNA
structure.  In this model, loop and stem regions are
mutually nested (\figref{tkfst}).  Both loops and stems have geometric length
distributions with parameters $\kappa_l$ and $\kappa_s$.  Single bases
are inserted and deleted in loops with rates $\lambda_l$ and $\mu_l$;
similarly, base-pairs are inserted and deleted in stems with rates
$\lambda_s$ and $\mu_s$.  Insertions of a new stem into an existing
loop sequence (or deletions of an existing stem) occur at the same
rate as single-base insertions (or deletions) and can model
large-scale structural changes (\figref{tkfst-mutations}).

@@

@@

Guided by our experience with the \emph{nanos} 3' TCE and tRNA, where
the reconstruction problem was made easy by the presence of a close
outgroup, we conducted a simulation study to investigate the
dependence of reconstruction accuracy on outgroup branch length.
(We here use the term ``outgroup'' loosely to denote the
variable-length branch in our three-taxon study,
where the other two branches are held at unit length.)

@@

These results correspond to a best-case scenario in which we
have a perfectly-parametrized evolutionary model and exactly-known descendant structures.
Recent results
demonstrating that inference of sequence alignment is
robust to mis-parametrized probabilistic models \cite{Lunter2007}
suggest that the same may be true for probabilistic inference of
ancestral structures.
The effect of imperfect knowledge of the ancestral structures has not been explored.

>>>>

\supptext{2}\ includes a complete formal algorithm for exact null-cycle elimination in SCFGs,
along with procedures for probabilistically restoring null cycles
to sampled parse trees and Inside-Outside expectation counts.
Informally, the essence of the algorithm is contained within the following two steps:
\begin{description}
\item[(i)] separating bifurcations
into those which have one or more empty children (and can therefore be
represented using transition or termination rules) and those that have
two nonempty children;
\item[(ii)] replacing all empty paths through null states with effective direct transitions between non-null states,
obtaining sum-over-paths probabilities by inverting the grammar's transition matrix.
\end{description}

@@

The TKF Structure Tree (TKFST) model is based on the
Thorne-Kishino-Felsenstein (TKF) model of the stochastic evolution of
primary sequences via indel events \cite{ThorneEtal91}.  In the
original TKF model, sequence evolves under a time-homogeneous linear
birth-death-immigration process \cite{Kendall1948}.  Single characters (``links'')
are inserted with rate $\lambda$ and deleted with rate $\mu$.  At
equilibrium, sequences obey a geometric length distribution with
parameter $\kappa$.  Although this model has flaws (e.g., it lacks
affine gap penalties, rate heterogeneity and context-dependent
mutation rates), it illustrates many of the key ideas used by more
sophisticated indel models, notably the possibility for systematic
derivation of pairwise alignment automata from first principles via
analysis of birth-death processes \cite{Feller71,ThorneEtal91}.

The TKF Structure Tree model is an extension of the TKF model to RNA
structure.  In this model, loop and stem regions are
mutually nested (\figref{tkfst}):
the parameter $\pi_l(S)$ determines the proportion of links within loop sequences that are nested stems,
and every stem sequence has a nested loop at the end.
Single bases are inserted and deleted in loops with rates $\lambda_l$ and $\mu_l$;
similarly, base-pairs are inserted and deleted in stems with rates
$\lambda_s$ and $\mu_s$.
Both loops and stems have geometric length
distributions with parameters $\kappa_l = \lambda_l / \mu_l$ and $\kappa_s = \lambda_s / \mu_s$.  
Insertions of a new stem into an existing
loop sequence (or deletions of an existing stem) occur at the same
rate as single-base insertions (or deletions) and can model
large-scale structural changes (\figref{tkfst-mutations}).

@@

A few other useful statistics for the TKFST model:
the expected number of links in a loop sequence is $K_l = \frac{\lambda_l}{\mu_l - \lambda_l}$
and in a stem sequence $K_s = \frac{\lambda_s}{\mu_s - \lambda_s}$.
Since $\pi_l(S)$ of the links in a loop sequence are nested stems,
and since each stem has twice as many nucleotides as it has links (since each link is a base {\em pair}),
the expected number of bases in a loop sequence is
\[
B_l = \frac{K_l (1-\pi_l(S)) + 2K_l \pi_l(S)} {1 - \pi_l(S) K_l}
\]
The expected number of bases in a stem sequence is
\[
B_s = 2K_s + B_l
\]
The expected number of bases that are created/removed when a loop-sequence link is inserted/deleted is
\[
D_l = (1-\pi_l(S)) + \pi_l(S) B_s
\]
The expected number of stems directly rooted in a given loop sequence is
$U = \pi_l(S) K_l$
and the expected number of stems directly rooted in, {\bf or indirectly descended from}, a given loop sequence is
$V = U / (1 - U)$
(note that this is also the expected total number of loop sequences indirectly descended from a given loop sequence).
Therefore, in the equilibrium structure,
the expected number of stems is $V$;
of loops, $V+1$;
of unpaired bases, $K_l (1 - \pi_l(S)) (V+1)$;
and of base-pairs, $K_s V$.
In a tree with total branch length $T$,
the expected number of single-base deletions is $\mu_l T K_l (1 - \pi_l(S)) (V+1)$;
of base-pair deletions, $\mu_s T K_s V$;
and of substructure deletions, $\mu_l T V$.

@@

Guided by our experience with the \emph{nanos} 3' TCE and tRNA, where
the reconstruction problem was made easy by the presence of a close
outgroup, we conducted a simulation study to investigate the
dependence of reconstruction accuracy on outgroup branch length,
and to investigate the performance of our reconstruction method (when simulating directly from the model)
compared to simpler reconstruction methods that ignore either structure or phylogeny.
(We here use the term ``outgroup'' loosely to denote the
variable-length branch in our three-taxon study,
where the other two branches are held at unit length.)

@@

These results correspond to a best-case scenario in which we
have a perfectly-parametrized evolutionary model and exactly-known descendant structures.
Recent results
demonstrating that inference of sequence alignment is
robust to mis-parametrized probabilistic models \cite{Lunter2007}
suggest that the same may be true for probabilistic inference of
ancestral structures.
%The effect of imperfect knowledge of the ancestral structures has not been explored.
% [commented out because the following tests DO address this in a roundabout way; what we mean is that indiegram has not been tested when leaf structures are unknown]

To compare \indiegram\ to simpler reconstruction methods, we also measured the alignment accuracy using various simpler statistical reconstruction techniques.
We evaluated three maximum likelihood inference algorithms that use simpler gap-aware phylogenetic models,
and one posterior decoding algorithm that uses a non-phylogenetic sum-of-pairs scoring scheme.
These results are plotted on \figref{simulations} alongside the \indiegram\ accuracies.
In detail, the reconstruction methods used were

\begin{description}
\item{TKF91 ML:} with the TKF91 model \cite{}, the \handel\ package \cite{HolmesBruno2001,Holmes2003,Holmes2007} was used to align the three extant sequences and reconstruct the ancestor.
The correct phylogenetic tree and branch lengths were supplied (as they were for the \indiegram\ benchmark).
The insertion, deletion and substitution rates for the TKF91 model were set equal to those of the loop submodel of TKFST.
\item{Long Indel ML:} with a single-event trajectory approximation to the long indel model \cite{MiklosLunterHolmes2004}, the \handel\ package was used to align the three extant sequences and reconstruct the ancestor.
The correct phylogenetic tree and branch lengths were supplied.
The deletion and substitution rates were set equal to those of the loop submodel of TKFST.
The mean indel length was set equal to $D_l$, the mean number of bases that are created/removed by an insertion/deletion in the loop submodel of TKFST;
the mean equilibrium sequence length (and thereby the insertion rate) was equal to $B_l$, the mean number of bases in TKFST at equilibrium
(\secname{A simple model of RNA structural evolution} has formulae for these quantities in terms of the TKFST rate parameters).
\item{Stemloc ML:} the \stemloc\ program was used to align the three sequences via single-linkage clustering with a Pair SCFG \cite{Holmes2005}.
The structures of the leaf sequences were provided, but not the phylogenetic branch lengths.
\item{Stemloc AMA:} the \stemloc\ program was used to align the three sequences in AMA mode, using posterior decoding to optimize a sum-over-pairs expected alignment accuracy \cite{BradleyPachterHolmes2008}.
The structures of the leaf sequences were provided, but not the phylogenetic branch lengths.
\end{description}

We did not measure the accuracy of the reconstructed ancestral structure in these tests (as we did in the \indiegram\ test),
since \handel\ does not consider structures, whereas \stemloc\ and \stemlocama\ do not provide ancestral reconstructions.

{\bf SUMMARY OF THESE RESULTS}
% TKF is less accurate than indiegram, although for short branch lengths it is almost identical; the TKF and indiegram accuracies diverge at longer branch lengths
% Long indel?
% Stemloc?
% StemlocAMA?

### 2009-06-04 s3.tex
renamed software appendix (Text S3) to Text S4; added new Text S3 for TKFST grammars; added a bunch of other stuff

\newpage
\section{Software}

All tools are available from \darturl\ as part of the \dart\ software package for sequence analysis.

\subsection{Automated grammar construction}
We implemented our model construction algorithm on the three-taxon star phylogeny
in a set of Perl scripts.
Given a singlet transducer modeling ancestral sequences and structures and
a branch transducer modeling structural evolution,
the scripts generate C++ code to create the corresponding jointly-normalized SCFG.
All possible models of structural evolution which can be represented by a Pair SCFG
are permitted as input to the scripts,
allowing for flexible and automated model design.

Given files describing the singlet and branch transducers,
including weights of all transitions which may be functions of evolutionary distance,
the package ComposedTreeTransducer::FourWayComposedTT can automatically generate
the state graph and transition matrix of a multi-sequence model of three extant sequences.
It removes the useless windback $\Snull$ states described in the paper and
introduces effective direct transitions caused by bifurcations with possibly-empty children.
The package ComposedTreeTransducer::TripletSCFG transforms a multi-sequence
model created by ComposedTreeTransducer::FourWayComposedTT into the 
corresponding jointly-normalized three-sequence SCFG
and generates C++ code to build the model.

Example singlet and branch transducers files are provided for a simple 
Pair HMM model, a simple Pair SCFG model and the full TKF Structure Tree model.

\subsection{Reconstruction of ancestral structures}
The program \indiegram\ can perform maximum-likelihood inference on the
three-sequence SCFGs automatically generated by the ComposedTreeTransducer::TripletSCFG package.
Complete or no structural information for the three extant sequences can be supplied
as input.

\subsection{Simulation of RNA family evolution}
The program that we wrote to simulate RNA structural alignments from the TKFST model, \evolsayer, is available as part of the same software distribution as \indiegram\ (the \dart\ package).
Another script (\animateevolsayer) can be used to make animations of the evolving RNA structures, in combination with the RNAplot program in the \viennarna\ package \cite{HofackerEtAl94}.

\bibliography{../latex-inputs/alignment,../latex-inputs/reconstruction,../latex-inputs/duplication,../latex-inputs/genomics,../latex-inputs/ncrna,../latex-inputs/url}

\end{document}

>>>>

\newpage
\section{Inference grammars for the TKFST model}
\seclabel{TKFST}

This section describes multi-sequence grammars that can be used to infer alignments, structures and ancestors under the
TKF Structure Tree (TKFST) model.

\input{tkfst.tex}

\newpage
\section{The TKF Structure Tree model as a transducer} \seclabel{tkfst}

We can cast the TKF Structure Tree (TKFST) model as a transducer.
This means we can automatically deduce rules like those shown for the TKFST model in \secref{TKFST}.

\subsection{The single-sequence TKFST model as a singlet transducer}

The states of the singlet transducer are shown in \tabref{tkfstsinglet}.
Allowed transitions between these states are shown in \secref{TKFST}.

\begin{table}[!ht]
  \centering
  \begin{tabular}{lrrrr}
    state & $\type$ & $\absorb$ & $\emit$ & $e(\,\bullet\,|\mathrm{TKFST})$ \\ \hline
    $L$ & $\Sstart$ \\
    $I_L$ & $\Sinsert$ & & $(x,\Tnull)$ & $p(x)$ \\
    \\
    $S$ & $\Sstart$ \\
    $I_S$ & $\Sinsert$ & & $(x,y)$ & $p(x,y)$ \\
    \\
    $B$ & $\Sinsert$ & & $(L\,S)$ & 1 \\
  \end{tabular}
  \caption{
    \tablabel{tkfstsinglet}
    States of the singlet transducer of the TKF Structure Tree model \cite{Holmes2004}.
    Singlet transducers can only have states of type $\Sstart$ or $\Sinsert$.
    This is the indiegram-style transducer equivalent of the singlet TKFST SCFG in \secref{TKFST}.
  }
\end{table}

\subsection{The two-sequence TKFST model as a branch transducer}

The states of the branch transducer are shown in \tabref{tkfstbranch}.
Allowed transitions between these states are shown in \secref{TKFST}.

\begin{table}[!ht]
  \centering
  \begin{tabular}{lrrrr}
    state & $\type$ & $\absorb$ & $\emit$ & $e(\,\bullet\,|\mathrm{TKFST})$ \\ \hline
    $L$ & $\Sstart$ \\
    $I_L$ & $\Sinsert$ & & $(u,\Tnull)$ & $p(u)$ \\
    $M_L$ & $\Smatch$ & $(x,\Tnull)$ & $(u,\Tnull)$ & $p(u|x)$ \\
    $D_L$ & $\Smatch$ & $(x,\Tnull)$ & $(\Tgap,\Tnull)$ & 1 \\
    $W_L$ & $\Swait$ \\
    \\
    $S$ & $\Sstart$ \\
    $I_S$ & $\Sinsert$ & & $(u,v)$ & $p(u,v)$ \\
    $M_S$ & $\Smatch$ & $(x,y)$ & $(u,v)$ & $p(u,v|x,y)$ \\
    $D_S$ & $\Smatch$ & $(x,y)$ & $(\Tgap,\Tgap)$ & 1 \\
    $W_S$ & $\Swait$ \\
    \\
    $B_i$ & $\Sinsert$ & & $(L_i\,S_i)$ & 1 \\
    $B$ & $\Smatch$ & $(L\,S)$ & $(L\,S)$ & 1 \\
    $B_p$ & $\Smatch$ & $(L\,S)$ & $(L\,\Send)$ & 1 \\
    $B_e$ & $\Smatch$ & $(L\,\Send)$ & $(L\,\Send)$ & 1 \\
  \end{tabular}
  \caption{\tablabel{tkfstbranch}
    States of the branch transducer of the TKF Structure Tree model \cite{Holmes2004}.
    States which have the same names as states of the singlet transducer in \tabref{tkfstsinglet}
    are the branch equivalent of the corresponding singlet states
    (e.g. a $\Smatch$ state might be the branch equivalent of an $\Sinsert$ state).
    States $L_i$ and $S_i$ are the $\Sstart$ states of a sub-model (not shown) identical
    in structure to the singlet transducer.  They are used to insert a new stem-loop structure.
    This is the indiegram-style transducer equivalent of the conditional pair TKFST SCFG in \secref{TKFST}.
  }
\end{table}

\subsection{The multi-sequence TKFST model as a composite transducer}

We can use the state graph construction algorithm described 
in the paper and detailed in Text S1 to create a model of the simultaneous evolution
of several sequences.

\begin{figure}[!htb]
  \centering
  \includegraphics [scale=0.4] {figs/parent.pdf}
  \caption{\figlabel{parent}
    A simple example of transducer composition to build a multi-sequence model of two extant sequences.
    An ancestral sequence $W$ evolves into two descendant sequences $X$ and $Y$.
    A singlet transducer (the horizontal gray box) emit ancestral sequence and structure
    and two branch transducers (the gray boxes labeled $\Delta T_X$ and $\Delta T_Y$)
    mutate it according to the specified multi-sequence model.
    Gray nodes correspond to observed data and white nodes unobserved data.
  }
\end{figure}

Consider the simple model shown in \figref{parent}.
The state of the multi-sequence model describing this model is a 3-vector $\bvec{a} = \left( a_1,\,a_2,\,a_3 \right)$,
where $a_1$ is the state of the singlet transducer generating the ancestral sequence $W$
and $a_2$ and $a_3$ are the states of the branch transducers evolving $W$ into
extant sequences $X$ and $Y$.

We can show some of the allowed transitions of this multi-sequence model.
The state of the branch transducer associated with the active node $n$ is shown in bold.

\paragraph{Stem creation.}

A stem is created at the root node $W$ (corresponding to a bifurcation in the singlet transducer $a_1$)
and survives in the sequence $X$ at node 2 but is deleted in the sequence $Y$ at node 3.

\begin{align}
  \nakedthreevec{1:}{2:}{3:} \quad
  \threevec{L}{L}{\mathbf{L}} &
  \to \threevec{\mathbf{L}}{W_L}{W_L}
  \to \threevec{B}{B}{B_p}
  \to \threevec{L}{L}{\mathbf{L}} \, \threevec{S}{\mathbf{S}}{\Send}
  \to \threevec{L}{L}{\mathbf{L}} \, \threevec{\mathbf{S}}{W_S}{\Send} \\
  & \to \threevec{\mathbf{L}}{W_L}{W_L} \, \threevec{\mathbf{S}}{W_S}{\Send}
\end{align}

\paragraph{Stem insertion.}

A stem sequence is inserted in sequence $X$ at node 2.

\begin{align}
  \nakedthreevec{1:}{2:}{3:} \quad
  \threevec{L}{L}{\mathbf{L}} \to \threevec{L}{\mathbf{L}}{W_L}
  \to \threevec{L}{B_i}{B}
  \to \threevec{L}{L}{\mathbf{L}} \, \threevec{\Send}{S}{\mathbf{S}}
  \to \threevec{\mathbf{L}}{W_L}{W_L} \, \threevec{\Send}{\mathbf{S}}{W_S}
\end{align}

\paragraph{Stem termination.}

All stem sequences are ended by (possibly empty) loop sequences.

\begin{align}
  \nakedthreevec{1:}{2:}{3:} \quad
  \threevec{\mathbf{S}}{W_S}{W_S}
  \to \threevec{B_e}{B_e}{B_e}
  \to \threevec{\Send}{\Send}{\Send} \, \threevec{L}{L}{\mathbf{L}}
  \to \threevec{\Send}{\Send}{\Send} \, \threevec{\Send}{\Send}{\Send}
\end{align}

The functions $\alpha(t)$, $\beta(t)$ and $\gamma(t)$ are parametrized by the insertion and deletion rates of the Structure Tree model.
They are defined for loop sequences as
\begin{align}
  \kappa_1 &= \lambda_1 / \mu_1 \\
  \alpha_1 &= \exp \left( -\mu_1 t \right) \\
  \beta_1 &= \frac{\lambda_1 \left( 1 - \exp \left((\lambda_1 - \mu_1) t \right) \right)}{\mu_1 - \lambda_1 \exp \left( (\lambda_1 - \mu_1) t \right) } \\
  \gamma_1 &= 1 - \frac{\mu_1 \left( 1 - \exp \left((\lambda_1 - \mu_1) t \right) \right)}{\left( 1 - \exp (- \mu_1 t) \right) \left(\mu_1 - \lambda_1 \exp \left( (\lambda_1 - \mu_1) t \right) \right) }
\end{align}
and similarly for stem sequences \cite{Holmes2004}.
